---
title: "homework1"
author: "Ali"
date: "Monday, February 09, 2015"
output:
  html_document:
    keep_md: yes
---

**Homework01**

Feel free to skip my housekeeping section and go straight to Question 01

Before starting the analysis, do some basic housekeeping:
- It is good practice to load libraries and define functions at the start of a script
- Pre-screen the data for missing values (we already know there are none)
- Have a look at the variables to get familiar with the experimental design, e.g. (str(design.txt))
- Read relevent background literature, e.g. (http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2527015/)
- look at the range and distribution of the data; does it make sense? Plotting datasets this large takes a lot of comput power; I like using summary(data.frame$variable) instead to check for balance, richness, etc.

**Housekeeping**
```{r}
#load libraries
library(car) #for recode
library(lattice) # for stripplot
data <- read.table("data.txt")
des <- read.table("design.txt")
newDat <- data.frame(gExp = as.vector(as.matrix(data)), sampleID = factor(colnames(data)), gene = factor(rownames(data)))
arDat <- suppressWarnings(data.frame(des, newDat))
summary(arDat$Treatment)
```

There is some discrepancy between the number of observations between Treatments.

```{r}
table(arDat$Treatment, arDat$time)
```

And the culprit is the 1 hr treatment. This might be due to the removal of missing values, and/or a problem with the microarray or experimental design. Let's expect the number of replicates:

```{r}
table(arDat$sampleID, arDat$time)
```

It looks like the intention was to have N = 6 for each time point; however, the 1 hr treatment only has N = 5. This will be important to keep in mind when performing statistics.

**Question 1a**

**Q** How many probes and how many samples?

**A** Simply take a look at the data.frame structure:
```{r}
str(arDat)
```

The number of probes is 22737, and the number of samples is 23.

**Question 1b**

**Q** What is the breakdown of samples for treatment and time?

**A** HAH! This question is addressing the same thing I discovered in my "housekeeping" section above. (Please refer to the above code for more details). There are 4 time points (1h, 2h, 4h, and 24h), with N = 6 each except for 1h, with N = 5. The design is slightly unbalanced.

**Question 1c**

**Q** Create a quantitative variable that represents the time at which cells were measured

**A** Use the 'recode' function in the 'cars' package.

```{r}
des$time <- recode(des$time, "'1_h' =1; '2_h'=2; '4_h'=4; '24_h'=24", as.factor.result = FALSE)
quantDat <- suppressWarnings(data.frame(des, newDat))
head(quantDat)
```

**Question 1d**

**Q** Create a plot showing the gene expression data for one probe and the averages for all possible combinations of Treatment and time

**A** I think the simplest solution would make a colour-coded stripplot. The line connects time categories in order and passes through the average for each treatment.

```{r}
#decide on a gene
set.seed(1)
(samp <- sample(1:nrow(arDat), size = 1))
#which gene did I get? 
head(arDat[samp, ])
specialGene = "1557628_s_at"
oneGene <- (subset(quantDat, subset = gene %in% specialGene))
#sanity check
head(oneGene[1:3, ])
#a stripplot function I wrote for another seminar:
makeStrip <- function(df) {
  stripplot(gExp ~ as.factor(time) | gene, df,
          group = Treatment, xlab = "Time after exposure (hrs)", jitter.data = FALSE, cex=1.5, lwd = 2,  
          auto.key = TRUE, type = c('p', 'a'))
}
makeStrip(oneGene)
```
** TO DO **

- Report averages for each treatment, time (the line passes through the average, but this is not explicit)
- format the figure more nicely, maybe in ggplot2 instead of lattice
