---
title: "seminar02b"
author: "Ali"
date: "Wednesday, January 21, 2015"
output:
  html_document:
    keep_md: yes
---

tosses = times the coin is flipped in one try
tries = toss events
p = probability of getting a head
turn coinflips into a function


```{r}
coinflips <- function(tosses, tries, p){
  tab <- matrix(runif(tosses * tries) > p,
              nrow = length(tosses), ncol=tries)
  y <- rowSums(tab)
  y.mean <- mean(y)
  prop <- (y.mean/tries)
  return(prop)
}
coinflips(1, 100, 0.5)
```

loop over many tries

B <- 1000
n <- 1
track <- matrix(nrow = B, ncol = n, dimnames = list(c(1:B),names(n)))
for (i in 1:B) {
  track[i,] <- coinflips(1, i, 0.5)
}
tail(track)
```{r}
B <- 1000
n <- 1
track <- matrix(nrow = B, ncol = n, dimnames = list(c(1:B),names(n)))
for (i in 1:B) {
  track[i,] <- coinflips(1, i, 0.5)
}
tail(track)
```
expect convergence around y = 0
make scatter plot

plot(track, xlab="flips", col="red")
legend("topright", c("p = 0.5"))

```{r, echo=FALSE}
plot(track, xlab="flips", col="red")
legend("topright", c("p = 0.5"))
```
or make a boxplot

boxplot(track)
legend("topright", c("p = 0.5"))

``{r, echo=FALSE}
boxplot(track)
legend("topright", c("p = 0.5"))
``
